---
title: "climate-drivers-markdown-seperated-models"
author: "Rachel Schattman"
date: "August 10, 2018"
output: html_document
---
<<<<<<< HEAD
---
title: "Climate Drivers Project @ SESYNC CYBERINFRASTRUCTURE COURSE"
author: "Rachel Schattman, Sarah Wiener, John Cobb, and Sarah Champion"
date: "July 2018, Annapolis MD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load packages
```{r}
library(MASS)
library(lme4)
library(rstanarm)
library("bayesplot")
library("ggplot2")
```

Data used in this project comes from three sources: (1) A survey of NRCS and FSA employees conducted by the USDA Climate Hubs in 2016/2017, (2) Crop idemnity payments made by FSA to farmers for weather-related crop loss between 2013-2016, and (3) mean and standard deviations of drought data over 1, 3, 5, 10, and 15 year periods.

```{r}
dataDR <- read.csv("C:\\Users\\rache\\Documents\\GitHub\\climate-drivers\\NRCS_FSAMergeDataset_w_PDSI2_7_28_18.csv", header = TRUE)
```
## Description of variables
### Dependent variable
The dependent variable is y_var, labled in our data frame as 'Concern_DryDrought'. This is a respondent concern about longer dry periods and drought, measured on a 4 point scale where 1 = not concerned, 2 = slightly concerned, 3 = concerned, 4 = very concerned

### Independent variables
For each regression model, we will use 2 independent variables. The first indedependent variable is 'PercLossDrought' which is the average percent loss due to drought in a state averaged over the years 2013-2016 and normalized for acres of farmland in the state.

Each model will also use of the following independent variables (timescales): 

*mean drought descriptions*

* 1 year (PDSI_MEAN_2016)
* 3 year (PDSI_MEAN_2014)
* 5 year (PDSI_MEAN_2010)
* 10 year (PDSI_MEAN_2007)
* 15 year (PDSI_MEAN_2002)

*standard deviation descriptions* 

* 1 year (PDSI_STD_2016)
* 3 year (PDSI_STD_2014)
* 5 year (PDSI_STD_2010)
* 10 year (PDSI_STD_2007)
* 15 year (PDSI_STD_2002)

We also include *Agency* as an independent variable. Each respondent to the survey is either an FSA or NRCS employee.


The spatial unit of analysis is 'stdiv', a combination of the statecode and the climate division within that state

## Preliminary data managmement
I assigned the dependent variable and ensure that variables are read as ordinal
```{r}
y_var_name <- "Concern_DryDrought"

dataDR$y_var <- dataDR[[y_var_name]]

dataDR$stdiv <- factor(dataDR$stdiv)
dataDR$Agency <- factor(dataDR$Agency)
dataDR$y_var <- factor(dataDR$y_var)
dataDR$PDSI_MEAN_2016 <- factor(dataDR$PDSI_MEAN_2016)
dataDR$PDSI_MEAN_2014 <- factor(dataDR$PDSI_MEAN_2014)
dataDR$PDSI_MEAN_2012 <- factor(dataDR$PDSI_MEAN_2012)
dataDR$PDSI_MEAN_2007 <- factor(dataDR$PDSI_MEAN_2007)
dataDR$PDSI_MEAN_2002 <- factor(dataDR$PDSI_MEAN_2002)
dataDR$PDSI_STD_2016 <- factor(dataDR$PDSI_STD_2016)
dataDR$PDSI_STD_2014 <- factor(dataDR$PDSI_STD_2014)
dataDR$PDSI_STD_2012 <- factor(dataDR$PDSI_STD_2012)
dataDR$PDSI_STD_2007 <- factor(dataDR$PDSI_STD_2007)
dataDR$PDSI_STD_2002 <- factor(dataDR$PDSI_STD_2002)
```

### Missing values
Identified and grouped the independent variables of interest. 
```{r}
x_var_clean <- c("PercLossDrought", 
                 "stdiv", 
                 "Agency", 
                 "PDSI_MEAN_2016",
                 "PDSI_MEAN_2014",
                 "PDSI_MEAN_2012",
                 "PDSI_MEAN_2007",
                 "PDSI_MEAN_2002",
                 "PDSI_STD_2016",
                 "PDSI_STD_2014",
                 "PDSI_STD_2012",
                 "PDSI_STD_2007",
                 "PDSI_STD_2002")

y_var_clean <- y_var_name

variables_used <- c(y_var_clean, x_var_clean)

```

I created a subset that removed observations with NAs. The number of observations is reduced to 4,760 observations, 15 variables.
```{r}
data_subset <- dataDR [,variables_used]
data_subset$y_var <- factor(data_subset[[y_var_name]])

data_subset <- na.omit(data_subset)

data_subset$y_var <- factor(data_subset[[y_var_name]])
```

### Now, onto the models!
My goal was to compare the fit of each of the 11 models listed below:
```{r}
list_models <- c("y_var ~ PercLossDrought + stdiv",
                 "y_var ~ PercLossDrought + PDSI_MEAN_2016", 
                 "y_var ~ PercLossDrought + PDSI_MEAN_2014",
                 "y_var ~ PercLossDrought + PDSI_MEAN_2012",
                 "y_var ~ PercLossDrought + PDSI_MEAN_2007",
                 "y_var ~ PercLossDrought + PDSI_MEAN_2002",
                 "y_var ~ PercLossDrought + PDSI_STD_2016",
                 "y_var ~ PercLossDrought + PDSI_STD_2014",
                 "y_var ~ PercLossDrought + PDSI_STD_2012",
                 "y_var ~ PercLossDrought + PDSI_STD_2007",
                 "y_var ~ PercLossDrought + PDSI_STD_2002")

```

Our original list of models looked like this. I am preserving it here in case we decide to use it again, but it will not run as code unless turned on.


list_models <- c("y_var ~ PercLossDrought + Agency",
                 "y_var ~ PercLossDrought + Agency + PDSI_MEAN_2016", 
                 "y_var ~ PercLossDrought + Agency + PDSI_MEAN_2014",
                 "y_var ~ PercLossDrought + Agency + PDSI_MEAN_2012",
                 "y_var ~ PercLossDrought + Agency + PDSI_MEAN_2007",
                 "y_var ~ PercLossDrought + Agency + PDSI_MEAN_2002",
                 "y_var ~ PercLossDrought + Agency + PDSI_STD_2016",
                 "y_var ~ PercLossDrought + Agency + PDSI_STD_2014",
                 "y_var ~ PercLossDrought + Agency + PDSI_STD_2012",
                 "y_var ~ PercLossDrought + Agency + PDSI_STD_2007",
                 "y_var ~ PercLossDrought + Agency + PDSI_STD_2002")


I set up the models to run in a future function
```{r}
n_model <- length(list_models)
list_mod <- vector("list", length=n_model)
```

I wanted to test several types of regression models, knowing that some form of ordinal model would  be most appropriate. I set up a loop with three types of models included. Turn model types *on* one at a time by removing the hashtag in front of it.
```{r}
# model_type <- "glm"
# model_type <- "ordinal"
model_type <- "bayes_stan"                        
```

Here is the loop
```{r}
for(i in 1:n_model){                #n_model = 1
  formula_model <- list_models[i]
  
  variables_used <- strsplit(formula_model,'\\+')
  variables_used <- unlist(strsplit(variables_used[[1]],'\\~'))
  variables_used <- unlist(lapply(variables_used,function(x){sub(" ","",x)}))
  variables_used <- unlist(lapply(variables_used,function(x){gsub(" ","",x)}))
  
  # variables_used
  
  data_subset <- dataDR [,variables_used]
  data_subset <- na.omit(data_subset)
  
  #data_subset$y_var <- factor(data_subset[[y_var_name]])
  
  if(model_type == 'glm'){                                   
    # generlized linear model from base R
    data_subset$y_var <- as.numeric(data_subset$y_var)
    
    mod <- try(glm(formula_model,
                   family = poisson,
                   data = data_subset))
  }
  if(model_type == "ordinal_model"){
    # ordinal regression model from MASS package
    mod <- try(polr(formula_model,
                    data = data_subset,
                    weights = rep(1, nrow(data_subset)),
                    Hess = TRUE,
                    start = rep(1,4)))
  }
  if (model_type == "bayes_stan"){
    # polr from the rstanarm package, for bayesian slope estimation 
    mod <- try(stan_polr(formula_model, 
                         data = data_subset, 
                         #prior = R2(0.25), 
                         prior = NULL,
                         prior_counts = dirichlet(1),
                         shape = NULL,
                         chains = 4, 
                         #cores = CORES, 
                         seed = 1234, 
                         iter = 200))
  }
  print(paste("i=",i))
  list_mod[[i]] <- mod   
  summary(list_mod[[1]]) 
  #save(mod,file= paste("C:\\Users\\rschattman\\Documents\\Research\\climate-drivers\\model",i,"output.rdata", sep ="")) # This save would be useful if you wanted to save each of the 11 models as their own file
}
```
```{r}
list_mod
```
###Tools for critiquing the model

```{r}
plot(loo(mod))
```

```{r}
loo1 <- (loo(mod, 
             save_psis = TRUE))
print(loo1)
```
```{r}
compare_models(mod)
```

```{r}
shinystan::launch_shinystan(mod)
```
### If you are going to save the model and use it later, the following will be useful:
Save all output to a single rdata file - this has been commented out to avoid rewriting my current saved file
```{r}
#save(list_mod,file="C:\\Users\\rschattman\\Documents\\Research\\climate-drivers\\modeloutput.rdata")
```

Load the models
```{r}
# for(i in 1:n_model){           #use this if you've saved multiple rdata files
# load(file= paste("C:\\Users\\rschattman\\Documents\\Research\\climate-drivers\\model",i,"output.rdata", sep =""))
# }

#use the code below if you've saved one rdata file with all the models in it

load(file=paste("C:\\Users\\rschattman\\Documents\\Research\\climate-drivers\\modeloutput.rdata", sep = ""))

```

Examine the characteristics of the rdata object
```{r}
str(list_mod)
str(summary(list_mod))
```

Display the coefficients - this code doesn't work yet.
```{r}

print(list_mod, digets = 3)
round(apply(rstan:: extract(list_mod$stanfit, pars = "drought") [[1]], 2, median), digits = 3)

loo(x, list_mod, cores = 1)
  save_psis = FALSE, K_threshold = NULL)

table1 <- data.frame (list_mod = 1, 
                    intercept = 1, 
                    intercept.se = 1, 
                    slope = 1, 
                    slope.se = 1, 
                    r.squared = 1, 
                    p.value = 1)
```

Extract coefficients from the rdata file and put them in a dataframe - also doesn't work yet
```{r}
for(i in 1:11){
  x<-rnorm(11)
  y<-rnorm(11)
  list_mod <- list_mod
  summary(list_mod)
}

for (i in 1:11){
  print(i)
}

table1[i,] <- c(i,
                summary(list_mod)[['coefficients']]['(Intercept)','Estimate'],
                     summary(list_mod)[['coefficients']]['(Intercept)','Std. Error'],
                summary(list_mod)[['coefficients']]['x','Estimate'],
                     summary(list_mod)[['coefficients']]['x','Std. Error'],
                summary(list_mod)[['r.squared']],
                     summary(list_mod)[['coefficients']]['x','Pr(>|t|)'])

write.csv(table1, file = 'table1.csv')
```
